A Note on Convergence Diagnostics for Multiple Imputation Algorithms

### 2019-12-12 Simulation/Writeup/NoteConvergence.tex
theoretical background restructured

The convergence properties of the MI algoritm in \pkg{mice} are investigated through model-based simulation. The results of this simulation study are guidelines for assessing convergence of MI algorithms. These guidelines will be implemented in an interactive evaluation tool for \pkg{mice}, `ShinyMICE', which is currently under development. All programming code used in this note is available from \href{https://github.com/gerkovink/shinyMice/simulation}{github.com/gerkovink/ShinyMICE/simulation}. 
%I will replicate Lacerda et al.'s simulation study on $\widehat{R}$ \citep{lace07}, and develop novel guidelines for assessing convergence. Ideally, I will integrate several diagnostics (e.g., $\widehat{R}$, \emph{auto-correlation}, and \emph{simulation error}) into a single summary indicator to flag non-convergence.

@@

Most convergence diagnostics target either of the two components. The stationarity component of convergence may be evaluated with auto-correlation \citep{scha97, gelm13}, numeric standard error \citep[which is commonly called `MC error']{gewe92}, or Raftery and Lewis's (\citeyear{raft91}) procedure to determine how many iterations are necessary to deminish the effect of trending within chains. The mixing component can be assessed with the potential scale reduction factor $\widehat{R}$ \citep[widely referred to as `the Gelman-Rubin statistic']{gelm92}. With an adapted version of $\widehat{R}$, proposed by \cite{veht19}, we may also evaluate the stationarity component of convergence. This would make $\widehat{R}$ a general convergence diagnostic. The application of $\widehat{R}$ to assess stationarity has not been thourougly investigated. Therefore, this study employs both $\widehat{R}$ and auto-correlation to investigate algorithmic convergence, as recommended by \cite[p.~898]{cowl96}.
%Therefore, the focus of this study is mainly on $\widehat{R}$. Stationarity will also be evaluated with auto-correlation as diagnostic, as recommended by \cite[p.~898]{cowl96}. 
Other convergence diagnostics are outside the scope of this study.\footnotemark 
\footnotetext{All of these methods evaluate the convergence of univariate scalar summaries (e.g., chain means or variances). These convergence diagnostics cannot diagnose convergence of multivariable statistics (i.e., relations between scalar summaries). \cite{buur18} proposed to implement multivariable evaluation by means of eigenvalue decomposition \cite{mack03}. This method is outside of the scope of the current study.}%  he eigenvector decomposition method proposed by McKay \citep{XYZ}.} These four methods, however, are outside of the scope of this study. Here, the focus is mainly on $\widehat{R}$ and auto-correlation.

%Diagnose non-convergence in the mixing sense: $\widehat{R}$. Diagnose non-convergence in the stability sense: auto-correlation (AC). Also, it is possible to compute MC errors, the Geweke statistic \textbf{\citep{XYZ}}, or to determine the appropriate number of iterations for convergence by employing \textbf{Raftery and Lewis's (1992)} method. All of these methods evaluate the convergence of univariate scalars of interest (e.g., chain mean or chain variance). The convergence of multivariable statistics (i.e. relations between such scalars) may be evaluated via the eigenvector decomposition method proposed by McKay \citep{XYZ}. These four methods, however, are outside of the scope of this study. Here, the focus is mainly on $\widehat{R}$ and auto-correlation.}

It is not known whether any convergence diagnostic is appropriate to diagnose convergence of \pkg{mice} algorithms. As by \cite[\S~6.5.2]{buur18}, the current practice is to visually inspect imputation chains for signs of non-mixing and non-stationarity. Visually assessing convergence, however, may be challenging to the untrained eye. \textbf{[This is problematic because]} There is no scientific consensus on the convergence properties of multiple imputation algorithms \citep{taka17}, and in some situations, the default techniques in \pkg{mice} might not converge to stable target distributions at all \citep{murr18}. %Convergence diagnostics for MI methods have not been systematically studied \citep{buur18}.  Moreover, 
%A fundamental assumption of MICE is convergence of the algorithm. MICE is a type of Markov chain Monte Carlo (MCMC) algorithm. In general, the validity of inference resulting from MCMC algorithms is threatened by non-convergence. Hence we use convergence diagnostics to flag non-convergence. But it is not known whether conventional convergence diagnostics for MCMC methods work on MI data.
Therefore, convergence of the \pkg{mice} algorithms should be monitored carefully. %the need for convergence diagnostics is extra pressing in the context of the MICE algorithm.

@@

%The potential scale reduction factor $\widehat{R}$ is obtained as:

High $\widehat{R}$ would indicate .... low $\widehat{R}$ values indicate ....
The conventionally acceptable threshold was $\widehat{R} < 1.2$ \cite{gelm92}. More recently, \cite{veht19} proposed a more stringent threshold: $\widehat{R} < 1.01$.

Non-recurrence can be evaluated with auto-correlation. Auto-correlation shows how dependent subsequent draws of an imputation chain are on the previous value. If there is a lot of dependence, draws at e.g. iteration five are significantly correlated with the value of the first draw. A high auto-correlation indicates dependence within chains. The magnitude of the AC can be interpreted qualitatively or quantitatively. Quantitative evaluation of the AC entails comparing the observed ACs to the critical values of a two-tailed 95\% confidence interval, divided by the square root of the number of iterations. %. critical value corresponding to the $z-$~value standard normal distribution with \textbf{XYZ} degrees of freedom.
%  AC_lim <- qnorm(.975) / sqrt(maxit)

\textbf{Convergence diagnostics applied to MI}

$\widehat{R}$ may not be an appropriate diagnostic to evaluate MI data because it assumes over-dispersed initial values. This means that the initial values of the $m$ imputation chains are `far away' from the distribution that the chains are converging to. In the \pkg{mice} algorithm, initial values are chosen randomly from the observed data. Therefore, we cannot be certain that the initial values are over-dispersed. %The initial parameter value is noodzakelijk. Daarom is het lastig om in MI that starten vanuit overdisp state.

% Starting values of MICE chains: "For each j, fill in starting imputations $Y^0_j$ by random draws from $Y^{obs}_j$." \cite[\S~4.5.2]{buur18} before for loops over variables and imputations, see \url{https://stefvanbuuren.name/fimd/sec-FCS.html#def:mice}

Without over-dispersed initial states, $\widehat{R}$ may falsely diagnose convergence \citep{broo98}. %would not be sensitive enough to flag non-convergence of MI algorithms: "if over-dispersion does not hold, $\sigma_+^2$ can be too low, which can lead to falsely diagnosing convergence." \cite[p~437]{broo98}. 
This suggests that $\widehat{R}$ would not be sensitive enough to flag non-convergence of MI algorithms. An empirical finding, however, shows that the opposite may be true: \cite{lace07} report $\widehat{R}$ values above the threshold of $\widehat{R} < 1.1$ after fifty iterations. %that $\widehat{R}$ will not be smaller than 1.1 before iteration number 50. Therefore, the aim is to replicate empirical finding Lacerda et al.

@@

The aim is to evaluate whether the imputation procedure has converged. The primary research interest is in determining whether $\widehat{R}$ is an appropriate convergence diagnostic, and if so, which level of stringency suits MI data. \textbf{Or if there is a discrepancy between the visual and numeric inspection. There is no baseline to compare the convergence diagnostics with. Therefore, we will only evaluate performance measures.}

This simulation study hypothesizes that $\widehat{R}$ will over-estimate non-convergence. \textbf{[Or use:] Hypothesis based on \cite{lace07} is that the conventional acceptable level of $\widehat{R}$ is too strict for MI data. We expect that the simulation diagnostics will indicate valid inference before $\widehat{R}$ will.}

@@

% \subsubsection{What is already implemented?}
% 
% - Trace-plots
% 
% \subsubsection{What is not yet implemented, but exists?}
% 
% - $\widehat{R}$, but too stringent (new) threshold, and assumption of over-dispersed initial values of imputation chains not met.
% % Potentially add something about updated (2019) version of $\widehat{R}$ and the new threshold of 1.01, see \citep{veht19}.
% 
% - Auto-correlation. Schafer (1997, p. 129) wrote on worst linear statistic. We could calculate the auto-correlation of that statistic to know that the algorithm converged elsewhere too. See autocorr function plot in SAS of worst linear function.
% % Worst linear function in SAS see \url{https://support.sas.com/documentation/cdl/en/statug/63033/HTML/default/viewer.htm#statug_mi_sect027.htm}.
% % Note: we're talking about missing data only, not the combined data (that autocorrelation is very high, as is the autocorrelation of deductively imputed values, like in the texp example). 
% % "Applications of MICE with lowly correlated data therefore inject a lot of noise into the system. Hence, the auto-correlation over t will be low, and convergence will be rapid, and in fact immediate if all variables are independent. Thus, the incorporation of noise into the imputed data has pleasant side-effect of speeding up convergence" \citep{buur18}, par. 4.5. 
% 
% - Sensitivity analysis: Run algorithm several times and compare results. 
% % "Imputers who do choose to use FCS should use flexible univariate models wherever possible and take care to assess apparent convergence of the algorithm, for example by computing traces of pooled estimates or other statistics and using standard MCMC diagnostics (Gelman et al., 2013, Chapter 11). It may also be helpful to examine the results of many independent runs of the algorithm with different initializations and to use random scans over the p variables to try to identify any convergence issues and mitigate possible order dependence" \cite[p.~19]{murr18}.
% 
% \subsubsection{What is not implemented, and NA?}
% 
% - $\widehat{R}$ threshold: Replicate simulation study and build a decision rule to solve the problem with $\widehat{R}$.
% % "The monitoring statistic was computed for mean monthly earnings at each iteration in chains of length k = 200. Since calculation of the statistic requires M parallel sequences, m = 5 such chains were constructed. This value of m was informed by the preferred choice given in the literature on multiple imputation. The monitoring statistic computed at each iteration is presented in Figures 15 to 17 for each of the three missingness mechanisms. The red vertical line denotes ten iterations" \cite[p.~49]{lace07}.
% 
% - Stability of the solution: Possibly use the slope of means over iterations too to see whether there is trending. Or apply PCA on the imputed data and if that (the eigenvalues?) stays the same we know that the means and variances are stable as well, see McKay (?). 
% 
% - MC error: MC error = SD/sqrt(number of iterations), where SD represents the variation across iterations. The MC error thus represents how much the means differ w.r.t. the iterations. MC error decreases as number of iterations increases. It should not be larger than 5\% of the sample standard deviation.

@@

\textbf{$\widehat{R}$}. The $\widehat{R}$ values for chain means and chain variances show equivalent trends, neither is there a clear difference between the four variables. Roughly speaking, we see that $\widehat{R}$ is higher for simulation conditions with a lower number of iterations, and approaches one for conditions with more iterations. Figure \ref{fig:conv} shows an initial dip, a small bump upwards, and then a steep decline. This decline \textbf{houdt aan} up-to iteration twenty, after which a more gradual decrease can be observed up-to iteration forty. The $\widehat{R}$ values are more or less stable after that. The \textbf{conventionally acceptable threshold for convergence, maximum $\widehat{R} < 1.2$,[CHECK IN INTRO]} is reached in the simulation condition with three iterations, and conditions with six iterations or more. The widely used threshold $\widehat{R} < 1.1$ is met in conditions with ten or more iterations. The most recent recommended threshold is not reached within the 100 iterations considered in this simulation study. $\widehat{R} > 1.01$ for all simulation conditions (see online appendix for the full table of results). \textbf{Check is computing r ht is still correct! Also say something about Rh below 1.}

\textbf{Auto-correlation.} All observed auto-correlations (aggregated over the 1000 repetitions) are negative or close to zero. There are some differences between variables, and the ACs of chain variances seem lower than the ACs of chain means. The general trend in figure \ref{fig:conv} is that ACs decrease steeply as the number of iterations increases.  \textbf{The lowest AC observed is %-0.735 
at four iterations. Simulation conditions with a higher number of iterations show a slow increase towards zero. Auto-correlations do not plateau completely within the range of simulation conditions considered in this study. It seems that ACs might become positive if more iterations would be allowed. The average absolute maximum ACs are negative up to 77 iterations for chain means, and for chain variances all iterations considered here have negative ACs.}

@@

From the simulation diagnostics, it appears that as little as three iterations might be sufficient to obtain unbiased, confidence valid estimates. Convergence diagnostics $\widehat{R}$ and auto-correlation, however, indicate that convergence may only be reached after twenty or even forty iterations.

\textbf{R hat below 1.}
$\widehat{R}$ could theoretically not be smaller than one, yet it occurred several times in this study (see online appendix XYZ). This can happen when the number of simulations is smaller than in `regular' MCMC processes [explain that fewer iterations is an advantage of MI, not a disadvantage compared to MCMC]. Therefore, the `$(n-1/n)$' [add equation number] correction factor can influence the estimated potential scale reduction factor. This downwards bias is in the opposite direction than expected: ``The mixture-of-sequences variance, $V$, should stabilize as a function of $n$. (Before convergence, we expect $\sigma^2$ to decrease with n, only increasing if the sequences explore a new area of parameter space, which would imply that the original sequences were not overdispersed for the particular scalar summary being monitored.)'' \cite[p~438]{broo98}.

>>>>

The convergence properties of the MI algoritm in \pkg{mice} are investigated through model-based simulation. The results of this simulation study are guidelines for assessing convergence of MI algorithms. These guidelines will be implemented in an interactive evaluation tool for \pkg{mice}, `ShinyMICE', which is currently under development. All programming code used in this note is available from \href{https://github.com/gerkovink/shinyMice/simulation}{github.com/gerkovink/ShinyMICE/simulation}. 
%I will replicate Lacerda et al.'s simulation study on $\widehat{R}$ \citep{lace07}, and develop novel guidelines for assessing convergence. Ideally, I will integrate several diagnostics (e.g., $\widehat{R}$, \emph{autocorrelation}, and \emph{simulation error}) into a single summary indicator to flag non-convergence.

@@

Most convergence diagnostics target either of the two components. The stationarity component of convergence may be evaluated with autocorrelation \citep[$AC$][]{scha97, gelm13}, numeric standard error \citep[or `MC error'][]{gewe92}, and Raftery and Lewis's (\citeyear{raft91}) procedure to determine how many iterations are necessary to deminish the effect of trending within chains. The mixing component can be assessed with the potential scale reduction factor $\widehat{R}$ \citep[widely referred to as `the Gelman-Rubin statistic']{gelm92}. With an adapted version of $\widehat{R}$, proposed by \cite{veht19}, we may also evaluate the stationarity component of convergence. This would make $\widehat{R}$ a general convergence diagnostic. The application of $\widehat{R}$ to assess stationarity has not been thourougly investigated. Therefore, this study employs both $\widehat{R}$ and autocorrelation to investigate algorithmic convergence, as recommended by \cite[p.~898]{cowl96}.
%Therefore, the focus of this study is mainly on $\widehat{R}$. Stationarity will also be evaluated with autocorrelation as diagnostic, as recommended by \cite[p.~898]{cowl96}. 
Other convergence diagnostics are outside the scope of this study.\footnotemark 
\footnotetext{All of these methods evaluate the convergence of univariate scalar summaries (e.g., chain means or variances). These convergence diagnostics cannot diagnose convergence of multivariable statistics (i.e., relations between scalar summaries). \cite{buur18} proposed to implement multivariable evaluation by means of eigenvalue decomposition \cite{mack03}. This method is outside of the scope of the current study.}%  he eigenvector decomposition method proposed by McKay \citep{XYZ}.} These four methods, however, are outside of the scope of this study. Here, the focus is mainly on $\widehat{R}$ and autocorrelation.

%Diagnose non-convergence in the mixing sense: $\widehat{R}$. Diagnose non-convergence in the stability sense: autocorrelation (AC). Also, it is possible to compute MC errors, the Geweke statistic \textbf{\citep{XYZ}}, or to determine the appropriate number of iterations for convergence by employing \textbf{Raftery and Lewis's (1992)} method. All of these methods evaluate the convergence of univariate scalars of interest (e.g., chain mean or chain variance). The convergence of multivariable statistics (i.e. relations between such scalars) may be evaluated via the eigenvector decomposition method proposed by McKay \citep{XYZ}. These four methods, however, are outside of the scope of this study. Here, the focus is mainly on $\widehat{R}$ and autocorrelation.}

@@

%The potential scale reduction factor $\widehat{R}$ is obtained as:
We can interpret $\widehat{R}$ as potential scale reduction factor, since it indicates by how much the variance of $\theta$ could be shrunken down if an infinite number of iterations per chain would be run \citep[$t \to \infty$;][]{gelm92}. High $\widehat{R}$-values would therefore indicate slow mixing of chains, e.g., because of dependence across chains on initial values. Once there is no dependence on initial values anymore, there is no added value of continuing iterations. The measure then diagnoses algorithmic convergence ($\widehat{R} = 1$). 
%If there is no added value compared to infinitely many iterations, the chains have converged. 
The conventionally acceptable threshold was $\widehat{R} < 1.2$ \cite{gelm92}. More recently, \cite{veht19} proposed a more stringent threshold: $\widehat{R} < 1.01$.

Following the same notation, autocorrelation is defined as the correlation between two subsequent $\theta$ values within the same chain. We only consider $AC$ at lag 1, i.e., between the $t^{th}$ and $t+1^{th}$ iteration of each chain: 
%
\begin{equation*}
AC = \left( \frac{T}{T-1} \right) \frac{\sum_{t=1}^{T-1}(\theta_t - \bar{\theta}^{(\cdot m)})(\theta_{t+1} - \bar{\theta}^{(\cdot m)})}{\sum_{t=1}^{T}(\theta_t - \bar{\theta}^{(\cdot m)})^2},
\end{equation*}
%
where $T$ is the total number of iterations per chain \citep[p.~147]{lync07}. We can interpret the sign of $AC$-values as measure of dependence within chains ($-1 \leq AC \leq 1$). Positive $AC$-values would indicate recurrence: similar $\theta$-values for subsequent iterations, which may lead to trending. Negative $AC$-values also indicate dependence, but do not threathen the stationarity component of convergence. Instead of recurrence, there would be a `seesaw' effect within chains: subsequent $\theta$-values would deviate in opposite directions from the chain average $\theta^{(\cdot m)})$, increasing the variance of $\theta$. As convergence diagnostic, the interest is in positive $AC$-values.\footnotemark 
\footnotetext{The magnitude of $AC$-values can also be evaluated statistically, but that is outside of this note's scope.}

%, where $\theta$ is a scalar summary of interest (i.e., chain means and chain variances).
%Non-recurrence can be evaluated with autocorrelation. autocorrelation shows how dependent subsequent draws of an imputation chain are on the previous value. If there is a lot of dependence, draws at e.g. iteration five are significantly correlated with the value of the first draw. 
%A high autocorrelation indicates dependence within chains. The magnitude of the AC can be interpreted qualitatively or quantitatively. Quantitative evaluation of the AC entails comparing the observed ACs to the critical values of a two-tailed 95\% confidence interval, divided by the square root of the number of iterations. %. critical value corresponding to the $z-$~value standard normal distribution with \textbf{XYZ} degrees of freedom.
%  AC_lim <- qnorm(.975) / sqrt(maxit)

It is not known whether $\widehat{R}$ and $AC$ are appropriate measures to diagnose convergence of MI algorithms. Assessing the stationarity component of convergence with $AC$ might be redundant, since high $AC$-values are implausible in MI procedures. That is, the randomness induced by the MI algorithm effectively mitigates the risk of dependency within chains. $AC$ would thus not be informative of the convergence of MI algorithms. $\widehat{R}$ may not be an appropriate diagnostic either, because it assumes over-dispersed initial values (i.e., the starting points of the chains are `far away' from the target distribution and each other). In \pkg{mice}, initial values of the algorithm are chosen randomly from the observed data. Therefore, we cannot be certain that the initial values are over-dispersed. %The initial parameter value is noodzakelijk. Daarom is het lastig om in MI that starten vanuit overdisp state. 
% Starting values of MICE chains: "For each j, fill in starting imputations $Y^0_j$ by random draws from $Y^{obs}_j$." \cite[\S~4.5.2]{buur18} before for loops over variables and imputations, see \url{https://stefvanbuuren.name/fimd/sec-FCS.html#def:mice}
Without over-dispersed initial states, $\widehat{R}$ may falsely diagnose convergence \citep{broo98}. %would not be sensitive enough to flag non-convergence of MI algorithms: "if over-dispersion does not hold, $\sigma_+^2$ can be too low, which can lead to falsely diagnosing convergence." \cite[p~437]{broo98}. 
This suggests that $\widehat{R}$ would not be sensitive enough to flag non-convergence of MI algorithms. An empirical finding, on the other hand, shows that $\widehat{R}$ can indicate non-convergence at unconventionally long chain lengths. While the default number of iterations in \pkg{mice} is five, $\widehat{R} > 1.1$ was observed after 50 iterations, and $\widehat{R} > 1.01$ upto 150 iterations. This suggests that $\widehat{R}$ would over-estimate non-convergence of MI algorithms.

% the conventional threshold $\widehat{R} < 1.1$ might be too stringent. \cite[pp.~49-50]{lace07} report
% 
% $\widehat{R}$ is sensitive enough to indicate non-convergence can  reported $\widehat{R}$-values  
% 
% An empirical finding, however, shows that the opposite may be true: \cite{lace07} report $\widehat{R}$-values above the threshold of $\widehat{R} < 1.1$ after fifty iterations. %that $\widehat{R}$ will not be smaller than 1.1 before iteration number 50. Therefore, the aim is to replicate empirical finding Lacerda et al. 
%

@@

This study evaluates whether $\widehat{R}$ and $AC$ can diagnose when the multiple imputation algorithm has converged. There is, however, no baseline measure available to evaluatate their performance against. The current practice of assessing convergence is to visually inspect imputation chains for signs of non-mixing and non-stationarity \citep[\S~6.5.2]{buur18}. Visually assessing convergence, however, may be challenging to the untrained eye. \textbf{[This is problematic because. MAYBE REPLACE THIS PART!]} There is no scientific consensus on the convergence properties of multiple imputation algorithms \citep{taka17}, and in some situations, the default techniques in \pkg{mice} might not converge to stable target distributions at all \citep{murr18}. %Convergence diagnostics for MI methods have not been systematically studied \citep{buur18}.  Moreover, 
%A fundamental assumption of MICE is convergence of the algorithm. MICE is a type of Markov chain Monte Carlo (MCMC) algorithm. In general, the validity of inference resulting from MCMC algorithms is threatened by non-convergence. Hence we use convergence diagnostics to flag non-convergence. But it is not known whether conventional convergence diagnostics for MCMC methods work on MI data.
Therefore, convergence of the MI algorithms in \pkg{mice} should be monitored carefully. %the need for convergence diagnostics is extra pressing in the context of the MICE algorithm.

%The primary research interest is in determining whether $\widehat{R}$ is an appropriate convergence diagnostic, and if so, which level of stringency suits MI data. \textbf{Or if there is a discrepancy between the visual and numeric inspection. There is no baseline to compare the convergence diagnostics with. Therefore, we will only evaluate performance measures.}

This simulation study hypothesizes that $\widehat{R}$ will over-estimate non-convergence. %\textbf{[Or use:] Hypothesis based on \cite{lace07} is that the conventional acceptable level of $\widehat{R}$ is too strict for MI data. We expect that the simulation diagnostics will indicate valid inference before $\widehat{R}$ will.}

@@

% \subsubsection{What is already implemented?}
% 
% - Trace-plots
% 
% \subsubsection{What is not yet implemented, but exists?}
% 
% - $\widehat{R}$, but too stringent (new) threshold, and assumption of over-dispersed initial values of imputation chains not met.
% % Potentially add something about updated (2019) version of $\widehat{R}$ and the new threshold of 1.01, see \citep{veht19}.
% 
% - autocorrelation. Schafer (1997, p. 129) wrote on worst linear statistic. We could calculate the autocorrelation of that statistic to know that the algorithm converged elsewhere too. See autocorr function plot in SAS of worst linear function.
% % Worst linear function in SAS see \url{https://support.sas.com/documentation/cdl/en/statug/63033/HTML/default/viewer.htm#statug_mi_sect027.htm}.
% % Note: we're talking about missing data only, not the combined data (that autocorrelation is very high, as is the autocorrelation of deductively imputed values, like in the texp example). 
% % "Applications of MICE with lowly correlated data therefore inject a lot of noise into the system. Hence, the autocorrelation over t will be low, and convergence will be rapid, and in fact immediate if all variables are independent. Thus, the incorporation of noise into the imputed data has pleasant side-effect of speeding up convergence" \citep{buur18}, par. 4.5. 
% 
% - Sensitivity analysis: Run algorithm several times and compare results. 
% % "Imputers who do choose to use FCS should use flexible univariate models wherever possible and take care to assess apparent convergence of the algorithm, for example by computing traces of pooled estimates or other statistics and using standard MCMC diagnostics (Gelman et al., 2013, Chapter 11). It may also be helpful to examine the results of many independent runs of the algorithm with different initializations and to use random scans over the p variables to try to identify any convergence issues and mitigate possible order dependence" \cite[p.~19]{murr18}.
% 
% \subsubsection{What is not implemented, and NA?}
% 
% - $\widehat{R}$ threshold: Replicate simulation study and build a decision rule to solve the problem with $\widehat{R}$.
% % "The monitoring statistic was computed for mean monthly earnings at each iteration in chains of length k = 200. Since calculation of the statistic requires M parallel sequences, m = 5 such chains were constructed. This value of m was informed by the preferred choice given in the literature on multiple imputation. The monitoring statistic computed at each iteration is presented in Figures 15 to 17 for each of the three missingness mechanisms. The red vertical line denotes ten iterations" \cite[p.~49]{lace07}.
% 
% - Stability of the solution: Possibly use the slope of means over iterations too to see whether there is trending. Or apply PCA on the imputed data and if that (the eigenvalues?) stays the same we know that the means and variances are stable as well, see McKay (?). 
% 
% - MC error: MC error = SD/sqrt(number of iterations), where SD represents the variation across iterations. The MC error thus represents how much the means differ w.r.t. the iterations. MC error decreases as number of iterations increases. It should not be larger than 5\% of the sample standard deviation.

@@

\textbf{$\widehat{R}$}. The $\widehat{R}$-values for chain means and chain variances show equivalent trends, neither is there a clear difference between the four variables. Roughly speaking, we see that $\widehat{R}$ is higher for simulation conditions with a lower number of iterations, and approaches one for conditions with more iterations. Figure \ref{fig:conv} shows an initial dip, a small bump upwards, and then a steep decline. This decline \textbf{houdt aan} up-to iteration twenty, after which a more gradual decrease can be observed up-to iteration forty. The $\widehat{R}$-values are more or less stable after that. The \textbf{conventionally acceptable threshold for convergence, maximum $\widehat{R} < 1.2$,[CHECK IN INTRO]} is reached in the simulation condition with three iterations, and conditions with six iterations or more. The widely used threshold $\widehat{R} < 1.1$ is met in conditions with ten or more iterations. The most recent recommended threshold is not reached within the 100 iterations considered in this simulation study. $\widehat{R} > 1.01$ for all simulation conditions (see online appendix for the full table of results). \textbf{Check is computing r ht is still correct! Also say something about Rh below 1.}

\textbf{autocorrelation.} All observed autocorrelations (aggregated over the 1000 repetitions) are negative or close to zero. There are some differences between variables, and the ACs of chain variances seem lower than the ACs of chain means. The general trend in figure \ref{fig:conv} is that ACs decrease steeply as the number of iterations increases.  \textbf{The lowest AC observed is %-0.735 
at four iterations. Simulation conditions with a higher number of iterations show a slow increase towards zero. autocorrelations do not plateau completely within the range of simulation conditions considered in this study. It seems that ACs might become positive if more iterations would be allowed. The average absolute maximum ACs are negative up to 77 iterations for chain means, and for chain variances all iterations considered here have negative ACs.}

@@

From the simulation diagnostics, it appears that as little as three iterations might be sufficient to obtain unbiased, confidence valid estimates. Convergence diagnostics $\widehat{R}$ and autocorrelation, however, indicate that convergence may only be reached after twenty or even forty iterations.

\textbf{R hat below 1.}
$\widehat{R}$ could theoretically not be smaller than one, yet it occurred several times in this study (see online appendix XYZ). This can happen when the number of simulations is smaller than in `regular' MCMC processes [explain that fewer iterations is an advantage of MI, not a disadvantage compared to MCMC]. Increasing Rhat values mean that the initial values were not appropriately over-dispersed \citep[p~438]{broo98}. %Therefore, the `$(n-1/n)$' [add equation number] correction factor can influence the estimated potential scale reduction factor. This downwards bias is in the opposite direction than expected: ``The mixture-of-sequences variance, $V$, should stabilize as a function of $n$. (Before convergence, we expect $\sigma^2$ to decrease with n, only increasing if the sequences explore a new area of parameter space, which would imply that the original sequences were not overdispersed for the particular scalar summary being monitored.)'' \cite[p~438]{broo98}.
